### 2019-03-19 paper-report.tex
init

\documentclass[11pt, letterpaper]{report}
\usepackage[margin=1in]{geometry}
\usepackage[utf8]{inputenc}
\usepackage{titlesec}

\titleformat{\chapter}[block]
{\normalfont\huge\bfseries}{\thechapter}{20pt}{}
\titlespacing*{\chapter}
{0pt}{20pt}{20pt}
\title{Paper Report "MacroBase: Prioritizing Attention in Fast Data"}
\author{Fayán Leonardo Pardo Ladino}
\date{March 2019}
\begin{document}
	\maketitle
	\chapter{Introduction}
	\begin{flushleft}
		In this report i am going to summarize and explain the paper "MacroBase: Prioritizing Attention in Fast Data". I am going to sum up main concept and description of MacroBase, the design behind it, the tests and results the researchers did on the solution, sum up their conclusions and finally I will relate it with the IBD course and expose my conclusions gathered after reading the paper.
	\end{flushleft}
\end{document}